---
title: CV
name: Mariano
surname: Viz
position: "Impact-Driven Conservationist | Environmental Consultant | Specialist in Impact Evaluation & Conservation Program Analysis"
address: "Bren School of Environmental Science & Management, University of California Santa Barbara (UCSB)"
phone: +1 805 868 303
email: "mviz@bren.ucsb.edu"
github: marianoviz
linkedin: www.linkedin.com/in/marianoviz
date: "`r format(Sys.time(), '%B %Y')`"
aboutme: "Impact-Driven Conservationist | Environmental Consultant | Specialist in Impact Evaluation & Conservation Program Analysis"
output: 
  vitae::markdowncv:
    theme: davewhipp
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```



## Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Santa Barbara, CA", "2022", "Bren School of Environmental Science & Management", "Master of Environmental Science and Management",
  "Buenos Aires, Argentina", "2014", "National University of La Plata (UNLP)", "Doctor of Veterinary Medicine",
  "Buenos Aires, Argentina", "2017", "Latin American Faculty of Social Sciences -FLACSO-", "Specialist Degree in Human Development",
  "Buenos Aires, Argentina", "2017", "Latin American Council of Social Sciences -CLACSO-", "Higher Diploma in Public Policies for Equality") %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

## Conservation Program Analyst Experience

```{r}

library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Environmental Markets Lab (UCSB),", "9/21-9/22", "Santa Barbara, CA", "Graduate Student Researcher",
  "Conservation International (CI)", "2/22-7/22", "Remote", "Sustainable Business Consultant",
  "World Wide Fund for Nature (WWF)", "6/21-9/21", "Remote", "Impact Evaluation Intern",
  "Conservation International (CI)", "6/21-9/21", "Galapagos Islands, Ecuador", "Sustainable Business Analyst Intern"
  ) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

## Skills & Affiliations

**Computer:** Microsoft Office (Word, Excel, PowerPoint, Access), Google Suite, RStudio, Git, ArcGIS, QGIS


**Languages:** Spanish (Native), English (Advanced Working Proficiency; TOEFL 109/120)

**Gap-year Travel (3/18–8/19):** Europe (England, Spain, France), Southeast Asia (Vietnam, Cambodia, Laos,
Thailand, Malaysia, Singapore, Indonesia), Australia

**Professional Affiliations:** Member of the Interdisciplinary Group for the Development of Family Farming
(GIDAF); Associate Member Wildlife Health Australia, Latin American Fisheries Fellowship.
