---
title: "Spatial data visualization (interactive map & chloropleth)"
description: |
  Here, I explore the spatial distribution of oil spill events in California in 2008 using some spatial data viz.
author:
  - name: Mariano Viz
    url: {}
date: 02-25-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(sf)
library(tmap)
library(gstat)
library(stars)
```

**Data source:** California Department of Fish and Wildlife (CDFW), Office of Spill Prevention and Response (OSPR). (2009). Oil Spill Incident Tracking, 2008. California Department of Fish and Wildlife. Available at: https://map.dfg.ca.gov/metadata/ds0394.html

## Summary

In this report, I explore the spatial distribution of oil spill events in California in 2008. First, I make an interactive map showing the location of inland and marine oil spill incidents. Then, I examine the distribution of inland oil spill events and make a static chloropleth map displaying counts of inland spill incidents by county. Data: CDFW (2009).  



```{r}
# Read in data, update col names, and check projection:


#Read in Oil Spill Incident Tracking data and update col names to lowercase:
oil_spill <- read_sf("Oil_Spill_Incident_Tracking_%5Bds394%5D-shp", layer = "Oil_Spill_Incident_Tracking_%5Bds394%5D") %>% 
  clean_names() %>% 
  rename(Location = inlandmari) #rename col showing spill location

# Check Oil Spill Incident Tracking data projection:
#st_crs(oil_spill) # WGS 84 (EPSG,3857)


# Read in the CA county data, update col names to lowercase, and keep only names:
ca_counties <- read_sf("ca_counties", layer = "CA_Counties_TIGER2016") %>% 
  clean_names() %>% 
  select(name) 

# Check the CA county projection
  #st_crs(ca_counties) # WGS 84 (EPSG,3857)
    #Same Projection as Oil Spill Incident Tracking data (no need to set or transform CRS)


```


## Exploratory interactive map of oil spill events


```{r}
# Check data with a quick ggplot:
#ggplot() +
  #geom_sf(data = ca_counties) +
  #geom_sf(data = oil_spill, aes(color = Location), size = 0.5)

#Interactive plot (tmap)
tmap_mode("view")

tm_shape(oil_spill) +
  tm_dots("Location") 

```

**Figure 1.** Oil spill events in California in 2008. Events are divided based on location. Green and yellow points indicate inland and marine event locations, respectively. Data: CDFW (2009). 



## Static chloropleth map showing counts of inland oil spill events by county


```{r}
#Only keep inland oil spills:
inland_oil_spill <- oil_spill %>% 
  filter(Location == "Inland")

#Spatial join of data:
ca_oil_spill <- ca_counties %>%
  st_join(inland_oil_spill)

#Counts of inland oil spills by county:
oil_spill_counts <- ca_oil_spill %>% 
  count(name)
#oil_spill_subset

#Static chloropleth map:
ggplot(data = oil_spill_counts) +
  geom_sf(aes(fill = n), color = "white", size = 0.1) +
  scale_fill_gradientn(colors = c("lightsalmon1","orangered1","red4")) +
  theme_minimal() +
  labs(fill = "Inland oil spills (2008)")

```

**Figure 2.** Counts of California’s inland oil spill events by county in 2008. County color, from salmon to dark red, reflects the number of inland oil spill incidents. Data: CDFW (2009).
















